---
title: "MPG Dataset Analysis"
author: "Stanley Prakash | 23200443"
format: html
editor: visual
---

# Assignment 1 :

# Introduction

The `mpg` dataset from the built-in R contains data on fuel economy for 38 popular car models. The variables include manufacturer, model, year, engine displacement, and fuel efficiency in miles per gallon (mpg).

# Data Summary

**Let's look into a brief statistics of the dataset:**

```{r echo=FALSE}
#| warning: false
library(dplyr)
library(datasets)
library(ggplot2)

# Loading the dataset
data("mpg")

# Calculating summary statistics
summary_stats <- mpg %>%
  summarise(
    Mean = mean(hwy),
    Median = median(hwy),
    SD = sd(hwy),
    Mean_1 = mean(cty),
    Median_1 = median(cty),
    SD_1 = sd(cty)
  )

# Creating a summary data frame
summary_1 <- data.frame(
  Statistic = c('Mean', 'Median', 'Standard Deviation'),
  Highway_MPG = c(summary_stats$Mean, summary_stats$Median, summary_stats$SD),
  Engine_Displacement = c(summary_stats$Mean_1, summary_stats$Median_1, summary_stats$SD_1)
)

# Printing summary statistics 
print(summary_1)

```

-   **Number of observations**: 234
-   **Number of variables**: 11
-   **Variables**:
    -   `manufacturer`: Car manufacturer
    -   `model`: Model of the car
    -   `year`: Year of manufacture
    -   `displ`: Engine displacement (liters)
    -   `hwy`: Highway miles per gallon
    -   `cty`: City miles per gallon
    -   `class`: Vehicle class (e.g., compact, SUV)

# Plot: Highway MPG vs. Engine Displacement

Below is a scatter plot showing the relationship between engine displacement (`displ`) and highway miles per gallon (`hwy`). Generally, as engine displacement increases, highway mpg decreases, indicating that larger engines tend to be less fuel-efficient.

```{r echo=FALSE}
# Creating the plot
plot <- ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  labs(title = "Highway MPG vs Engine Displacement",
       x = "Engine Displacement (liters)",
       y = "Highway Miles per Gallon",
       color = "Car Class") +
  theme_minimal()

# Displaying the plot
print(plot)

```

As observed, there is an inverse relationship between engine displacement and highway mpg. This trend is common as larger engines typically consume more fuel.

# 

# Conclusion

The plot effectively illustrates the negative correlation between engine displacement and highway miles per gallon. This insight can be useful for consumers looking to purchase fuel-efficient vehicles and for manufacturers aiming to improve fuel economy in their car models.

## 
